---
title: "SOK1005_Assignment_3"
format: pdf
editor: visual
---

```{r}
rm(list=ls())
# we first start loading relevant packages and libraries to scrap data and perform the analysis  
library(tidyverse)
library(rvest)  # for web scraping 
library(janitor) # for data cleaning 
library(lubridate) #to work with date
library(quantmod) 
library(stringr)
```

## Task 1

```{r}
url <- "https://www.motor.no/aktuelt/motors-store-vintertest-av-rekkevidde-pa-elbiler/217132"

table <- url %>%
  read_html() %>% 
  html_nodes("table") %>% 
  html_table() 
head(table[[1]])

df_1 <- table[[1]]
df_2 <- table[[2]]
```

df_1\$STOPP=as.numeric(levels(df_1\$STOPP))\[df_1\$STOPP\]

df_1\$WLTP=as.numeric(levels(df_1\$WLTP))\[df_1\$WLTP\]

```{r}
df_1 <- subset(df_1) %>%
  rename("M_0_-10" = X1, WLTP = X2, STOPP = X3, Avvik = X4)

df_1 <- df_1[-1,]
df_1 <- df_1[-19,]
df_1 <- df_1[-25,]
```

```{r}
df_1$STOPP <- str_replace(df_1$STOPP, " km", "")

df_1[c('WLTP', 'kWh')] <- str_split_fixed(df_1$WLTP, '/', 2)

df_1$WLTP <- str_replace(df_1$WLTP, " km", "")

df_1$STOPP=as.numeric(df_1$STOPP)
df_1$WLTP=as.numeric(df_1$WLTP)
```

```{r}
df_1 %>% 
  ggplot(aes(x=WLTP, y=STOPP)) +
  geom_point()+ 
  geom_abline() +
  xlim(200,600) +
  ylim(200,600) +
  labs(title = "Oppgitt rekkevidde vs testet rekkevidde",
       subtitle = "Testet ved 0 til -10 Celcius",
       y="Testet rekkevidde",
       x="Oppgitt rekkevidde",
       caption = "Kilde: https://www.motor.no/aktuelt/motors-store-vintertest-av-rekkevidde-pa-elbiler/217132")
```

As we can see in the figure, none of the cars lives up to the expectations of their advertised range. (Under the conditions 0 to -10 Celcius)

If they lived up to their advertised range, they would be at the 45 degree line, and if they surpassed the expectations they would be above the 45 degree line.

## Task 2

```{r}
lm(STOPP ~ WLTP, data = df_1)

#geom_smooth(method = lm)
```

The WLTP seems to be about 86.71% of the expected range on average.

```{r}
df_1 %>% 
  ggplot(aes(x=WLTP, y=STOPP)) +
  geom_point()+ 
  geom_abline() +
  xlim(200,600) +
  ylim(200,600) +
  geom_smooth(method = lm)+
  labs(title = "Oppgitt rekkevidde vs testet rekkevidde",
       subtitle = "Testet ved 0 til -10 Celcius",
       y="Testet rekkevidde",
       x="Oppgitt rekkevidde",
       caption = "Kilde: https://www.motor.no/aktuelt/motors-store-vintertest-av-rekkevidde-pa-elbiler/217132")
```
