---
title: "Sok-1005 Exam"
format: html
editor: visual
---

# Prosjektoppgave Sok-1005, Vår 2023

```{r}
rm(list=ls()) 
library(tidyverse)
library(rvest)
library(janitor)
library(lubridate)
library(quantmod) 
library(stringr)
library(foreign)
library(mefa)
```

```{r}
# Add the csv I made into a dataframe in oppgave 1

df = read.csv('https://raw.githubusercontent.com/TrymWS/Sok-1005/main/Assignments/Prosjektoppgave/mov_cust_demo.csv')

# Change weeks from 17-68 to 1-52

df <- df %>%
  mutate(across(week, ~  . - 16))
```

## Oppgave 2

I understood weekly reporting as a report containing 52 weekly points, since the data wrangling requested to merge with a dataset that only contained weekly datapoints

```{r}
# Choose store 73

Store_73 <- df %>% filter(store %in% (73))

Store_73 <- Store_73 %>% mutate(G_D_M = (grocery + dairy + meat)/1000)
```

```{r}
Store_73 %>%
  ggplot(aes(week, soda_sales)) +
  geom_point(col="black") +
  geom_line(col="blue") +
  
  scale_x_continuous(breaks=c(0, 5,10,15,20,25,30,35,40,45,50))+
  
  labs(y = "Sales", x = "Week", title = "Coca Cola sales", 
subtitle = "Sales per week i Dollars") +
  
  theme_minimal()
```

```{r}
Store_73 %>%
  ggplot(aes(week, grocery)) +
  geom_point(col="black") +
  geom_line(col="blue") +
  
  scale_x_continuous(breaks=c(0, 5,10,15,20,25,30,35,40,45,50))+
  
  labs(y = "Sales", x = "Week", title = "Grocery sales", 
subtitle = "Sales per week in Dollars") +
  
  theme_minimal()
```

```{r}
Store_73 %>%
  ggplot(aes(week, dairy)) +
  geom_point(col="black") +
  geom_line(col="blue") +
  
  scale_x_continuous(breaks=c(0, 5,10,15,20,25,30,35,40,45,50))+
  
  labs(y = "Sales", x = "Week", title = "Dairy sales", 
subtitle = "Sales per week in Dollars") +
  
  theme_minimal()
```

```{r}
Store_73 %>%
  ggplot(aes(week, meat)) +
  geom_point(col="black") +
  geom_line(col="blue") +
  
  scale_x_continuous(breaks=c(0, 5,10,15,20,25,30,35,40,45,50))+
  
  labs(y = "Sales", x = "Week", title = "Meat sales", 
subtitle = "Sales per week in Dollars") +
  
  theme_minimal()
```

```{r}
Store_73 %>%
  ggplot(aes(week, G_D_M)) +
  geom_point(col="black") +
  geom_line(col="blue") +
  
  scale_x_continuous(breaks=c(0, 5,10,15,20,25,30,35,40,45,50))+
  
  labs(y = "Sales x $1,000", x = "Week", title = "Grocery, Dairy and Meat sales", 
subtitle = "Sales per week in 1,000 Dollars") +
  
  theme_minimal()
```

From the figures, we can see that groceries, dairy and meat sales correspond quite a lot with eachother, which is no big surprise.

We can also see that the sale of the soda brands I chose in total does not seem to correspond with daily food consumption.

There are several weeks that are distinctly more active in sales than others, for both consumption groups. Often close to special events / holidays.

So sales of the various goods should be included in such a report.

## Oppgave 3

For the monthly report I will combine 4 weeks to represent a month, and look at that.

```{r}
df_sum <- df %>% 
  group_by(week) %>% 
    select_if(is.numeric) %>%
    summarise(across(c(2:55),sum),
            .groups = 'drop') %>%
  as.data.frame()

df_month <- df_sum %>% filter(week %in% (20:23))

df_month <- df_month %>% 
    select_if(is.numeric) %>%
    summarise(across(c(1:55),sum),
            .groups = 'drop') %>%
  as.data.frame()

 # summarise(across(where(is.numeric), ~ list(summary(.))))

df_month_1 <- df_month %>%
  select(1:54)

df_month_1 <- df_month_1[-1] %>% 
            t() %>% 
            as.data.frame() %>% 
            setNames(df_month_1[,1])

df_month_1 <- cbind(var = rownames(df_month_1), df_month_1)
rownames(df_month_1) <- 1:nrow(df_month_1)

colnames(df_month_1)[colnames(df_month_1) == "86"] ="value"

# Remove coupons, customer count, misc and profit, left with sales
# Removed grocery since it dominated the chart
df_month_1 <- df_month_1[-c(9,2,4,6,8,13,14,17,19,21,25,26,27,29,35,38,43,48,52,53,54),]

df_month_1 <- df_month_1 %>% mutate(value = value/1000)
```

```{r}
df_month_1 %>%
  ggplot(aes(value, var)) +
  geom_col(fill="black", col="grey") +
  #scale_x_continuous(breaks=c(0))+
  
  labs(y = "Goods", x = "x $1,000", title = "Sales numbers", 
subtitle = "Sales numbers for the various goods, measured in thousands of dollars.") +
  
  theme_minimal()
```

The sales of individual goods for that month is an important thing to include.

Perhaps also the prior month, so you can evaluate the difference. Or last years same month, so you can see what the year over year change is.

## Oppgave 4

```{r}
df_weeks <- df_sum
df_weeks <- df_weeks %>% mutate(grocery = grocery/1000)

df_weeks %>%
  ggplot(aes(week, grocery)) +
  geom_point(col="black") +
  geom_line(col="blue") +
  
  scale_x_continuous(breaks=c(0, 5,10,15,20,25,30,35,40,45,50))+
  
  labs(y = "Sales x $1,000", x = "Week", title = "Grocery sales", 
subtitle = "Sales per week in x1,000 Dollars") +
  
  theme_minimal()
```

```{r}
df_weeks <- df_weeks %>% mutate(custcoun = custcoun/1000)

df_weeks %>%
  ggplot(aes(week, custcoun)) +
  geom_point(col="black") +
  geom_line(col="blue") +
  
  scale_x_continuous(breaks=c(0, 5,10,15,20,25,30,35,40,45,50))+
  
  labs(y = "Customer count x 1,000", x = "Week", title = "Customer count", 
subtitle = "Amount of customers in total per week over the year") +
  
  theme_minimal()
```

Grocery sales and customer count can be a good thing to have when evaluating opening a new store. Aggregating all the sales should also be done.

```{r}
df_year <- df %>% 
  group_by(store) %>% 
    summarise(across(c(3:55),sum),
            .groups = 'drop') %>%
  as.data.frame()

df_year <- df_year %>% mutate(custcoun = custcoun/1000)

df_year <- cbind(var = rownames(df_year), df_year)
rownames(df_year) <- 1:nrow(df_year)

df_year %>%
  ggplot(aes(var,custcoun)) +
  #geom_point(col="black") +
  #geom_line(col="blue") +
  geom_col(fill="black", col="grey") +

  
  #scale_y_continuous(breaks=c(0, 5,10,15,20,25,30,35,40,45,50))+
  
  labs(y = "Customer count x 1,000", x = "Reset store number", title = "Customer count", 
subtitle = "Amount of customers in total per week over the year") +
  
  theme_minimal()
```

Customer count based on which stores they visit can be helpful in locating new locations.

One could assume that if you find multiple store, within a certain area, that is outperforming the mean, it could be a good idea to locate another store in that vicinity
